---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# decido

The goal of decido is to demonstrate a minimal R binding to the 
Mapbox library [earcut.hpp](https://github.com/mapbox/earcut.hpp). 

This is to augment the JS version available in [rearcut](https://github.com/hypertidy/rearcut.git) (also an R wrapper
of the JS version by Mapbox). 

## Installation

You can install decido from with:

``` r
devtools::install_github("hypertidy/decido")
```

## Development

Immediate needs: 

* allow input of holes
* wrap around complex types of multiple polygons, etc. 
* tests!


## Example

This is a basic example of triangulating a single-ring polygon. The
output is a vector of triplet indices defining each triangle. 


```{r example}
library(decido)
x <- c(0, 0, 0.75, 1, 0.5, 0.8, 0.69)
y <- c(0, 1, 1, 0.8, 0.7, 0.6, 0)
earcut(x, y)


plot(x, y)
polygon(x, y, lwd = 10)
apply(matrix(earcut(x, y), nrow = 3), 2, function(ind) polygon(cbind(x, y)[ind + 1, ], col = "firebrick"))
```


```{r}
library(oz)
oz_ring <- oz::ozRegion(states = FALSE)
ring <- oz_ring$lines[[6]]
plot(ring, pch = ".", asp = 1/cos(42 * pi/180))
indices <- earcut(ring$x, ring$y)
xy <- cbind(ring$x, ring$y)
apply(matrix(indices, nrow = 3), 2, function(ind) polygon(xy[ind + 1, ]))
```

## Performance

Compare timing of C++ versus JS implementations. 


```R
 rbenchmark::benchmark(rearcut::earcut(cbind(ring$x, ring$y)), decido::earcut(ring$x, ring$y))
                                    test replications elapsed relative
#2     decido::earcut(ring$x, ring$y)          100   0.064    1.000
#1 rearcut::earcut(cbind(ring$x, ring$y))          100   4.147   64.797
```
## Beware!

Only single ring polygons are supported currently. No checking is done
on input coordinates, so use at your own risk!  
