---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis-CI Build Status](http://badges.herokuapp.com/travis/hypertidy/decido?branch=master&env=BUILD_NAME=trusty_release&label=linux)](https://travis-ci.org/hypertidy/decido)
[![Build Status](http://badges.herokuapp.com/travis/hypertidy/decido?branch=master&env=BUILD_NAME=osx_release&label=osx)](https://travis-ci.org/hypertidy/decido) 
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hypertidy/decido?branch=master&svg=true)](https://ci.appveyor.com/project/mdsumner/decido) 
[![Coverage status](https://codecov.io/gh/hypertidy/decido/branch/master/graph/badge.svg)](https://codecov.io/github/hypertidy/decido?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidync)](https://cran.r-project.org/package=tidync)


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# decido

The goal of decido is to provide an R binding to the 
Mapbox library [earcut.hpp](https://github.com/mapbox/earcut.hpp) for constrained polygon triangulation. Decido is aimed at package developers
at the moment, there are not high-level classes or objects but the earcut
functionality can be easily used in higher-level tools or just used directly. 

Ear cutting (or ear clipping) applies constrained triangulation by 
successively 'cutting' triangles from a polygon defined by path/s. Holes are supported, the earcut library works with 
single-island-with-holes polygons, analogous to the POLYGON type in simple features. 

This augments the Javascript version available in [rearcut](https://github.com/hypertidy/rearcut.git) (also an R wrapper of the JS version by Mapbox). Only very minimal comparison  testing has yet been done to [compare these implementations](http://rpubs.com/cyclemumner/364247). 


## Installation

You can install decido from GitHub with the following. You will need
the [set of development tools](https://www.rstudio.com/products/rpackages/devtools/) for building source packages with binary code. 


```R
## install.packages("devtools")
devtools::install_github("hypertidy/decido", build_vignettes = TRUE)
```

## Example

This is a basic example of triangulating a single-ring polygon. The
output is a vector of triplet indices defining each triangle. 


```{r example}
library(decido)
x <- c(0, 0, 0.75, 1, 0.5, 0.8, 0.69)
y <- c(0, 1, 1, 0.8, 0.7, 0.6, 0)
earcut(x, y)
```

See the [documentation](https://hypertidy.github.io/decido/) and [vignette](https://hypertidy.github.io/decido/articles/decido.html) for more. 

Open the getting started vignette. 

```R
vignette("decido", package = "decido")
```

## Development

Immediate needs: 

* wrap around complex types of multiple polygons, etc. 
* more doc and tests!

This is motivated by the topology aspirations of  [hypertidy/silicate](https://github.com/hypertidy/silicate). We need
tools for decomposing shape data into primitives for analysis and visualization. Decomposition into graph types is already well supported and exercised, but triangulations of paths versus triangulations from edges are two key facilities needed for greater control. 


## Beware!

This is in active development and the functions will change. 

No checking is done on input coordinates, so use at your own risk!  


## Other implementations

Ear clipping (or ear cutting) is also available in the [rgl](https://CRAN.r-project.org/package=rgl) function `triangulate` (implemented in R), and in the [lawn](https://CRAN.r-project.org/package=lawn) function `lawn_tesselate` (implemented via the Mapbox Javascript library earcut).  In rgl the function also classifies input coordinates according to their nesting, a necessary first step if the relaionship between holes and islands is not known.  The `INLA` package has some
kind of constraint-based triangulation, but I don't yet know the details. 

In comparison to path-based ear-clipping, other libraries 'Triangle' and 'CGAL' provide edge-based *mostly Delaunay* triangulation. The Triangle library is available in the R package [RTriangle](https://CRAN.r-project.org/package=RTriangle). Experimental implementations binding CGAL are in [rcgal](https://github.com/s-u/rcgal) and [laridae](https://github.com/hypertidy/laridae). 


Do you know of others? Let me know!  Triangulation is common across 
many R packages, but constrained algorithms are pretty rare (it's hard).  There are many Delaunay and other non-constrained implementations in many packages, and I'm compiling a list of those as well. OTOH there's rgeos, sf, deldir, geometry, tripack, spatstat, akima, and several mesh-related packages Rvcg, meshsimp, icosa, webglobe ...



---

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).  By participating in this project you agree to abide by its terms.
