---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis-CI Build Status](http://badges.herokuapp.com/travis/hypertidy/decido?branch=master&env=BUILD_NAME=trusty_release&label=linux)](https://travis-ci.org/hypertidy/decido)
[![Build Status](http://badges.herokuapp.com/travis/hypertidy/decido?branch=master&env=BUILD_NAME=osx_release&label=osx)](https://travis-ci.org/hypertidy/decido) 
[![AppVeyor Build Status](https://ci.appveyor.com/api/projects/status/github/hypertidy/decido?branch=master&svg=true)](https://ci.appveyor.com/project/mdsumner/decido) 
[![Coverage status](https://codecov.io/gh/hypertidy/decido/branch/master/graph/badge.svg)](https://codecov.io/github/hypertidy/decido?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/tidync)](https://cran.r-project.org/package=tidync)


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# decido

The goal of decido is to provide an R binding to the 
Mapbox library [earcut.hpp](https://github.com/mapbox/earcut.hpp) for constrained polygon triangulation.  

Ear cutting (or ear clipping) applies constrained triangulation by 
successively 'cutting' triangles from a polygon defined by path/s. Holes are supported, the earcut library works with 
single-island-with-holes form polygons, analogous to the POLYGON type in simple features. 

This augments the Javascript version available in [rearcut](https://github.com/hypertidy/rearcut.git) (also an R wrapper of the JS version by Mapbox). No comparison or testing has yet been done to compare these implementations. 

In comparison to path-based ear-clipping, other libraries 'Triangle' and 'CGAL' provide edge-based *mostly Delaunay* triangulation. The Triangle library is available in the R package [RTriangle](https://CRAN.r-project.org/package=RTriangle). Experimental implementations binding CGAL are in [rcgal](https://github.com/s-u/rcgal) and [laridae](https://github.com/hypertidy/laridae). 

## Installation

You can install decido from with:

``` r
devtools::install_github("hypertidy/decido")
```

## Development

Immediate needs: 

* wrap around complex types of multiple polygons, etc. 
* tests!


See [rearcut](https://github.com/hypertidy/rearcut.git) for some approaches to the hole encoding. It's not hard, just rife with options. At the moment it seems best to stick close to how the JS earcut encodes them, and this translates very simply in C++. 

## Example

This is a basic example of triangulating a single-ring polygon. The
output is a vector of triplet indices defining each triangle. 


```{r example}
library(decido)
x <- c(0, 0, 0.75, 1, 0.5, 0.8, 0.69)
y <- c(0, 1, 1, 0.8, 0.7, 0.6, 0)
earcut(x, y)
```

See the documentation and vignette for more. 



## Beware!

This is in active development and the functions will change. 

No checking is done on input coordinates, so use at your own risk!  


---

Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).  By participating in this project you agree to abide by its terms.
